---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```

```{r load-data}
brfss2013 %>%
  summarise(count = n())
```

* * *

## Part 1: Data
The Behavioral Risk Factor Surveillance System (BRFSS) data is collected from U.S. residents regarding their health-related risk behaviors, chronic health conditions, and use of preventive services via telephone surveys.
This system started in 1984 and covered all 50 states as well as the District of Columbia and three U.S. territories. It completed more than 400,000 adult interviews each year.

* * *

## Part 2: Research questions

**Research quesion 1:**
How does joint pain severity affect sleep time of US citizens?

**Research quesion 2:**
Does smoking increase the likelihood of getting heart attack or cancer?

**Research quesion 3:**


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**
Compare percentages of people with inadequate sleep time (< 6 hours)
within people who reported activity limited due to joint pain and everyone in this survey who reported sleep time (not NA).

```{r}
joint_sleep <- brfss2013 %>%
  filter(!is.na(sleptim1) & sleptim1 < 15 & !is.na(lmtjoin3) & lmtjoin3 == "Yes") %>%
  mutate(inadq_slp = sleptim1 < 6)

joint_sleep %>%
  summarise(std_slp = sd(sleptim1), mean_slp = mean(sleptim1))

```

```{r}

joint_sleep %>%
  select(sleptim1, inadq_slp, joinpain)

```

```{r}
joint_sleep %>%
  summarise(inadq_slp_rate = sum(inadq_slp == TRUE)/n())
```


```{r}
all_sleep <- brfss2013 %>%
  filter(!is.na(sleptim1) & sleptim1 < 15) %>%
  mutate(inadq_slp = sleptim1 < 6)

all_sleep %>%
  summarise(inadq_slp_rate = sum(inadq_slp == TRUE)/n())
```
Almost 20% of people with joint symptoms have inadequate sleep time.
Only 10.8% general public have inadequate sleep time.


**Research quesion 2:**
smoke100: Smoked At Least 100 Cigarettes (Yes, No)
chcocncr: (Ever Told) You Had Any Other Types Of Cancer? (Yes, No)
cvdinfr4: Ever Diagnosed With Heart Attack? (Yes, No)

```{r}
sm_htcr <- brfss2013 %>%
  filter(!is.na(cvdinfr4),!is.na(chcocncr), smoke100 == "Yes") %>%
  select(cvdinfr4,smoke100,chcocncr)

sm_htcr <- sm_htcr %>%
  mutate(smk_ca_ht = ifelse(cvdinfr4== "Yes" | chcocncr == "Yes", "Yes", "No"))

sm_group <- sm_htcr %>%
  group_by(smk_ca_ht) %>%
  summarise(count=n())

ggplot(data=sm_group,aes(x=smk_ca_ht,y=count))+geom_bar(stat='identity')+xlab("Has either heart attack or cancer")+ylab("Number of US Citizens")
```

```{r}
sm_htcr %>%
  summarise(ht_or_ca = sum(smk_ca_ht == "Yes")/n())
```
About 18.4% people who Smoked At Least 100 Cigarettes had either heart attack or cancer
About 11.3% people Not smoked At Least 100 Cigarettes had either heart attack or cancer

```{r}
nosm_htcr <- brfss2013 %>%
  filter(!is.na(cvdinfr4),!is.na(chcocncr), smoke100 == "No") %>%
  select(cvdinfr4,smoke100,chcocncr)
nosm_htcr <- nosm_htcr %>%
  mutate(ca_ht = ifelse(cvdinfr4== "Yes" | chcocncr == "Yes", "Yes", "No"))
nosm_htcr %>%
  summarise(ht_or_ca = sum(ca_ht == "Yes")/n())
```

**Research quesion 3:**
qlmentl2: How Many Days Depressed In Past 30 Days
exerhmm2: Minutes Or Hours Walking, Running, Jogging, Or Swimming
exerhmm1: Minutes Or Hours Walking, Running, Jogging, Or Swimming
exeroft1: How Many Times Walking, Running, Jogging, Or Swimming
exerany2: Exercise In Past 30 Days

```{r}
ex_dep <- brfss2013 %>%
  filter(!is.na(qlmentl2),!is.na(exerhmm2), !is.na(exeroft1)) %>%
  select(qlmentl2, exerhmm2, exeroft1)
ex_dep %>%
  summarize(meam = mean(exeroft1), std = sd(exeroft1))

ex_dep <- ex_dep %>%
  mutate(enough_ex = ifelse(exeroft1 > 130, "Yes", "No"))


```
There were less people who felt depressed more than five days over the past 30 days in group who exercise (Walking, Running, Jogging, Or Swimming) more than 130 times/month than those who exercise < 130 times/month

```{r}
ex_30 <- ex_dep %>%
  group_by(enough_ex) %>%
  summarise(dep = sum(qlmentl2 > 5))

ggplot(data=ex_30,aes(x = enough_ex, y = dep)) + geom_bar(stat='identity')+xlab("walk, jog > 130 times/month")+ylab("more than 5 days feel depressed")
```